---
title: "DATA 624 Homework 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---


```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(fpp2)
library(forcats)
library(readxl)
library(seasonal)
library(forecast)
```


# Problem 6.2
The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

**a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?**
There appears to be seasonaility and an upward trend cycle 
```{r}
plastics <- plastics
autoplot(plastics)
```

**b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.**
```{r}
multiplicative_decomposition <- decompose(plastics, type = "multiplicative")
autoplot(multiplicative_decomposition)
```
**c. Do the results support the graphical interpretation from part a?**
Yes the results from the muliplicative decomposition do match the graphical interpretation from part a, we can see that there is an upward trend and there is a strong seasonality component.

**d. Compute and plot the seasonally adjusted data.**
```{r}
autoplot(plastics, series="Data") +
  autolayer(trendcycle(multiplicative_decomposition), series="Trend") +
  autolayer(seasadj(multiplicative_decomposition), series="Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales in Thousands") +
  ggtitle("Plastic Product Sales") +
  scale_colour_manual(values=c("gray","blue","red"), breaks=c("Data","Seasonally Adjusted","Trend"))
```

**e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?**
```{r}
outlier_plastics <- plastics
outlier_plastics[20] <- outlier_plastics[20] + 500
multiplicative_decomposition_new <- decompose(outlier_plastics, type = "multiplicative")

autoplot(outlier_plastics, series = "Data") +
  autolayer(trendcycle(multiplicative_decomposition_new), series = "Trend") +
  autolayer(seasadj(multiplicative_decomposition_new), series = "Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales in Thousands") +
  ggtitle("Plastic Product Sales") +
  scale_color_manual(values=c("gray", "blue", "red"), breaks=c("Data", "Seasonally Adjusted", "Trend"))
```
The outlier slighty affects the trend but dramatically effects the seasonally adjusted data.


**Does it make any difference if the outlier is near the end rather than in the middle of the time series?**
The outlier seems to have a slighly stronger impact on the trend component when it is at the end of the series

```{r}
outlier_plastics <- plastics
outlier_plastics[50] <- outlier_plastics[50] + 500
multiplicative_decomposition_new <- decompose(outlier_plastics, type = "multiplicative")

autoplot(outlier_plastics, series = "Data") +
  autolayer(trendcycle(multiplicative_decomposition_new), series = "Trend") +
  autolayer(seasadj(multiplicative_decomposition_new), series = "Seasonally Adjusted") +
  xlab("Year") + ylab("Monthly Sales in Thousands") +
  ggtitle("Plastic Product Sales") +
  scale_color_manual(values=c("gray", "blue", "red"), breaks=c("Data", "Seasonally Adjusted", "Trend"))
```

# Problem 6.3 
**Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?**

There appears to be some spikes in the remainder graphs that may be an indicator of the prescence of outliers 

```{r}
retaildata <- read_excel("retail.xlsx", skip = 1)
retail <- ts(retaildata[, "A3349337W"], frequency = 12, start = c(1982, 4))
x11_retail <- seas(retail, x11="")
autoplot(x11_retail) +
  ggtitle("X11 Decomposition of Retail Sales Data")
```

