---
title: "DATA 624 Homework 9"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r, echo = T, results = 'hide', include=FALSE}
library(caret)
library(pls)
library(tidyverse)
library(AppliedPredictiveModeling)
library(corrplot)
library(kableExtra)

```


# Question 7.2

**Friedman (1991) introduced several benchmark data sets created by simulation. On of these simulations used the following nonlinear equations to create data:**

\[y = 10 sin(\pi x_1x_2) + 20(x_3- 0.5)^2 + 10x_4 + 5x_5 + N(0, \sigma^2)\]

**where the x values are random variables uniformly distributed between [0,1] (there are also 5 other non-informative variables also created in the simulation). The package mlbench contains a function called mlbench.friedman1 that simulates these data:**

```{r}
library(mlbench)
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
## We convert the 'x' data from a matrix to a data frame
## One reason is that this will five the columns names.
trainingData$x <- data.frame(trainingData$x)
## Look at the data using
featurePlot(trainingData$x, trainingData$y)
## or other methods.

## This creates a list with a vector 'y' and a matrix
## of predictors 'x'.  Also simulate a large test set to
## estimate the true error rate with good precision:
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)

```

**Tune several models on these data. For example:**

## KNN MODEL
```{r}
library(caret)
knnModel <- train(x = trainingData$x, 
                  y = trainingData$y,
                  method = "knn",
                  preProcess = c("center", "scale"),
                  tuneLength = 10)
knnModel

```

```{r}
knnPred <- predict(knnModel, newdata = testData$x)
## The function 'postResample' can be used to get the test set
## performance values
postResample(pred = knnPred, obs = testData$y)
```

**Which models appear to give the best performance? Does MARS select the informative predictors (those named X1-X5)?**

## MARS MODEL

```{r}
library(earth)
MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
MARS_model <- train(x = trainingData$x, 
                  y = trainingData$y,
                  method = "earth",
                  tuneGrid = MARS_grid,
                  preProcess = c("center", "scale"),
                  tuneLength = 10)
MARS_model
```
### Look at the Perfomance of the MARS Model
```{r}
MARS_predictions <- predict(MARS_model, newdata = testData$x)
postResample(pred = MARS_predictions, obs = testData$y)
```

### Look at Variable Importance 

```{r}
varImp(MARS_model)
```

## SVM MODEL

```{r}
library(kernlab)
SVM_model <- train(x = trainingData$x,
                   y = trainingData$y,
                   method = "svmRadial",
                   preProcess = c("center", "scale"),
                   tuneLength = 10,
                   trControl = trainControl(method = "cv"))
SVM_model
```

### Results of the SVM Model and the Most Important Variables for SVM
```{r}
SVM_predictions <- predict(SVM_model, newdata = testData$x)
postResample(pred = SVM_predictions, obs = testData$y)
varImp(SVM_model)
```

## Neural Network Model 

```{r}
nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
nnet_maxnwts <- 5 * (ncol(trainingData$x) + 1) + 5 + 1
nnet_model <- train(x = trainingData$x,
                    y = trainingData$y,
                    method = "avNNet",
                    preProcess = c("center", "scale"),
                    tuneGrid = nnet_grid,
                    trControl = trainControl(method = "cv"),
                    linout = TRUE,
                    trace = FALSE,
                    MaxNWts = nnet_maxnwts,
                    maxit = 500)
nnet_model
```

### Results of the Neural Net Model and the Most Important Variables for the Neural Net
```{r}
nnet_predictions <- predict(nnet_model, newdata = testData$x)
postResample(pred = nnet_predictions, obs = testData$y)
varImp(nnet_model)
```

## Random Forest Model

```{r}

#10 folds repeat 3 times
control <- trainControl(method='repeatedcv', 
                        number=10, 
                        repeats=3)
#Metric compare model is Accuracy
metric <- "Accuracy"
set.seed(123)
#Number randomely variable selected is mtry
mtry <- sqrt(ncol(trainingData$x))
tunegrid <- expand.grid(.mtry=c(1:10))
rf_default <- train(x = trainingData$x,
                    y = trainingData$y,
                      method='rf', 
                      metric='RMSE', 
                      tuneGrid=tunegrid, 
                      trControl=control)
rf_default
```


```{r}
rf_predictions <- predict(rf_default, newdata = testData$x)
postResample(pred = rf_predictions , obs = testData$y)

```





## Look at the Results of the Models
```{r}
results <- data.frame(t(postResample(pred = knnPred, obs = testData$y))) %>% 
  mutate("Model" = "KNN")

results <- data.frame(t(postResample(pred = MARS_predictions, obs = testData$y))) %>%
  mutate("Model"= "MARS") %>%
  bind_rows(results)

results <- data.frame(t(postResample(pred = SVM_predictions, obs = testData$y))) %>%
  mutate("Model"= "SVM") %>%
  bind_rows(results)

results <- data.frame(t(postResample(pred = nnet_predictions, obs = testData$y))) %>%
  mutate("Model"= "Neural Network") %>%
  bind_rows(results)


results <- data.frame(t(postResample(pred = rf_predictions, obs = testData$y))) %>%
  mutate("Model"= "Random Forest") %>%
  bind_rows(results)

results %>%
  select(Model, RMSE, Rsquared, MAE) %>%
  arrange(RMSE) 


rm(results)
```

**The MARS model was the top performing model**


# Question 7.5

**Exercise 6.3 describes data for a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and training several nonlinear regression models.**

## Pre-Processing from last Homework

```{r}
library(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)
library(RANN)


knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
df <- predict(knn_model, ChemicalManufacturingProcess)


df <- df %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))

in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

```


## (a) Which nonlinear regression model gives the optimal resampling and test set performance?


### Partial Least squares
```{r}
pls_model <- train(
  Yield ~ ., data = train_df, method = "pls",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

pls_model

pls_predictions <- predict(pls_model, test_df)

results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "PLS")
```

```{r}
pls_predictions <- predict(pls_model, test_df)

results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "PLS")
```

### KNN Model

```{r}
knn_model <- train(
  Yield ~ ., data = train_df, method = "knn",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
knn_model

knn_predictions <- predict(knn_model, test_df)

results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "KNN") %>% rbind(results)

ggplot(knn_model, highlight = TRUE) + 
  labs(title = paste0("Tuning profile: ", knn_model$modelInfo$label))

```

### MARS Model

```{r}
MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)

MARS_model <- train(
  Yield ~ ., data = train_df, method = "earth",
  tuneGrid = MARS_grid,
  # If the following lines are uncommented, it throws an error
  #center = TRUE,
  #scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
MARS_model

MARS_predictions <- predict(MARS_model, test_df)

results <- data.frame(t(postResample(pred = MARS_predictions, obs=test_df$Yield))) %>%
  mutate("Model"= "MARS") %>% rbind(results)

ggplot(MARS_model, highlight = TRUE) + 
  labs(title = paste0("Tuning profile: ", MARS_model$modelInfo$label))
# Variable Importance
ggplot(varImp(MARS_model), top = 20) + 
  labs(title = paste0("Variable importance: ", MARS_model$modelInfo$label))
```

### Random Forest Model

```{r}
#10 folds repeat 3 times
control <- trainControl(method='repeatedcv', 
                        number=10, 
                        repeats=3)
#Metric compare model is Accuracy
metric <- "Accuracy"
set.seed(123)
#Number randomely variable selected is mtry
mtry <- sqrt(ncol(trainingData$x))
tunegrid <- expand.grid(.mtry=c(1:10))
rf_default <- train(Yield ~ ., data = train_df,
                      method='rf', 
                      metric='RMSE', 
                      tuneGrid=tunegrid, 
                      trControl=control)
rf_default

RF_predictions <- predict(rf_default, test_df)

results <- data.frame(t(postResample(pred = RF_predictions, obs=test_df$Yield))) %>%
  mutate("Model"= "RF") %>% rbind(results)


ggplot(knn_model, highlight = TRUE) + 
  labs(title = paste0("Tuning profile: ", knn_model$modelInfo$label))
# Variable Importance
ggplot(varImp(knn_model), top = 20) + 
  labs(title = paste0("Variable importance: ", knn_model$modelInfo$label))

```

### Neural Net Model

```{r}
nnetGrid <- expand.grid(decay = c(0, 0.01, 0.1), size = 1:10, bag = FALSE)
prep <- c("center", "scale", "nzv")
ctrl <- trainControl(method = "cv", number = 10)
nnetModel <- train(Yield ~ ., data = train_df,
                   preProcess = prep, 
                   trControl = ctrl, 
                   method = "avNNet", 
                   linout = TRUE, trace = FALSE, 
                   tuneGrid = nnetGrid)
nnetModel

NN_predictions <- predict(nnetModel, test_df)

results <- data.frame(t(postResample(pred = NN_predictions, obs=test_df$Yield))) %>%
  mutate("Model"= "Neural Net") %>% rbind(results)

ggplot(nnetModel, highlight = TRUE) + 
  labs(title = paste0("Tuning profile: ", nnetModel$modelInfo$label))

```

### K Nearest Neighbors

```{r}
knn_model <- train(
  Yield ~ ., data = train_df, method = "knn",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
knn_model

ggplot(knnModel, highlight = TRUE) + 
  labs(title = paste0("Tuning profile: ", knnModel$modelInfo$label))

knn_predictions <- predict(knn_model, test_df)

results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "KNN") %>% rbind(results)
```


## Results Summary
THe Random Forest Function seemed to be the highest performing algorithm in this case
```{r}
results %>%
  select(Model, RMSE, Rsquared, MAE) %>%
  arrange(RMSE) 
```


## Part B

**Which predictors are most important in the optimal nonlinear regression model? Do either the biological or process variables dominate the list? How do the top ten important predictors compare to the top ten predictors from the optimal linear model?**

```{r}
# Variable Importance
ggplot(varImp(knn_model), top = 20) + 
  labs(title = paste0("Variable importance: ", knn_model$modelInfo$label))

# Variable Importance
ggplot(varImp(nnetModel), top = 20) + 
  labs(title = paste0("Variable importance: ", nnetModel$modelInfo$label))


# Variable Importance
ggplot(varImp(pls_model), top = 20) + 
  labs(title = paste0("Variable importance: ", pls_model$modelInfo$label))

# Variable Importance
ggplot(varImp(MARS_model), top = 20) + 
  labs(title = paste0("Variable importance: ", MARS_model$modelInfo$label))

```

## Part C

**Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and their relationship with yield?**

The variables below all represent features that are strongly correlated with the target variable and they all showed up relatively high in variable importance for the different algorithms
```{r}
ggplot(train_df, aes(BiologicalMaterial12, Yield)) +
  geom_point()

ggplot(train_df, aes(BiologicalMaterial06, Yield)) +
  geom_point()

ggplot(train_df, aes(BiologicalMaterial03, Yield)) +
  geom_point()

ggplot(train_df, aes(ManufacturingProcess13, Yield)) +
  geom_point()

ggplot(train_df, aes(ManufacturingProcess17, Yield)) +
  geom_point()


ggplot(train_df, aes(ManufacturingProcess32, Yield)) +
  geom_point()



```


